Title: Moh hassan
========================================================

# Reproducible Research 
# Project 1
# Getting data from the web

# Set working dir - optional

```{r echo=FALSE}
setwd('~/RMIT15.071x/RS') 
```

```{r dataLoad}
# Download the dataset into workspace (wd), if it is not avalaible in local 
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
fileZip <- "dataset.zip"
if (!file.exists(fileZip)) download.file(fileUrl, fileZip, method = "auto")
# unzip the dataset into wd - dir 
if(file.exists(fileZip)){unzip(fileZip, exdir="./data")}
# variable that specifies the data path

# Read file
  activity <- read.csv("./data/activity.csv")
```

  # What is mean total number of steps taken per day?
  # For this part of the assignment, you can ignore the missing values in the dataset.
  # 1. Make a histogram of the total number of steps taken each day
  #activity$steps[is.na(activity$steps)] <- 0
  # or 
  #sum(activity$steps, igonre.na = TRUE)
  # aggreate by date
  
 
```{r}
 df <- aggregate( steps~ date, data=activity, FUN=sum)
  hist(df$steps, xlab='steps per day', main='Histogram of Steps per day for 2 months',     breaks=10)
  mean(df$steps)
  median(df$steps)
```

# What is the average daily activity pattern?
# 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
# and the average number of steps taken, averaged across all days (y-axis)
```{r}
  avg_daily <- aggregate( .~ interval, data=activity, FUN=sum) 
  plot(avg_daily$interval, avg_daily$steps,xlab="5 min interval", ylab="log(steps)", type="l")
```

# 2. Which 5-minute interval, on average across all the days in the dataset, 
#    contains the maximum number of steps?
```{r}
 avg_daily[avg_daily$steps==max(avg_daily$steps),]
```


# Imputing missing values
# 1. Calculate and report the total number of missing values in the dataset 
# (i.e. the total number of rows with NAs)
```{r}
  colSums(is.na(activity))
  sapply(activity, function(x) sum(is.na(x)))
```



```{r}
activity.complete <- activity
  # replace all steps with NA's with the mnean of steps (37,8)
  activity.complete$steps[is.na(activity.complete$steps)] <- mean(activity$steps, na.rm=TRUE)/2
  
  # Sum the number of steps per day over 2 months
  # 3. complete data 
  df.complete <- aggregate( steps~ date, data=activity.complete, FUN=sum)
  # 4 histogram 
  hist(df.complete$steps, xlab='steps per day', main='Histogram of Steps per day for 2 months', breaks=10)
  
  mean(df.complete$steps)
  median(df.complete$steps)
  #  Mean and media have higher value when replacing missing steps values
  #  with  avg steps / 2 to normalize steps data.
  # The impact results in having skewed data for missing values on a given date.
  
  # ---------------------------------------------------------------------------
  # Weekdays and Weekends differences
  # 1. Create a new factor variable for weekday and weekend

  df.complete.wd  <- aggregate( .~ interval, data=activity.complete, FUN=sum) 
  df.complete.wd$weekday <- weekdays(as.Date(df.complete.wd$date)) 
  
  weekend <- c('Saturday','Sunday')
  df.complete.weekend <- df.complete.wd[df.complete.wd$weekday %in% weekend,]
  
  # panel plot
  par(mfrow = c(2,1)) # 2 rows and 1 column
  

  plot(df.complete.wd$interval, df.complete.wd$steps,xlab="5 min interval", ylab="log(steps)", type="l")
  plot(avg_daily$interval, avg_daily$steps,xlab="5 min interval", ylab="log(steps)", type="l")

```

